### 2018-02-09 phd_1/0_main.tex
before unite to one file

\documentclass[10pt]{article}
\usepackage{femj_ru}
\usepackage{algorithmicx}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\captionsetup[algorithm]{labelformat=empty}
\usepackage[noend]{algpseudocode}
\usepackage{subfig}
\usepackage{caption}
\graphicspath{{img/}}

\begin{document}
\Pages(1--8)
\def\Im{\mathop{\mathrm{Im}}\nolimits}
\summary Mesenev~P.\,R., Chebotarev~A.\,Yu. \author
Boundary inverse problem for conductive-radiative equations of heat transfer\title
Problem of re-isteblishment of reflection properties of part of the boundary for stationary model of radiative-conductive heat transfer in a three0dimensional domain is considered. Presented an algorithm for solving the problem, which effectiveness is demonstrated by numerical experiments.
\keywords{functional equations, theta functions, Weierstrass sigma-function, nonlinear sequences.}
%Boundary inverse problem for complex equations of heat transfer\title
%Problem of re-esteblishment of relfeсtion properties of boundary part for stationary model of radiative-conductive heat transfer in a three-dimensional domain is considered. Presented an algorithm for solving the problem, which effectiveness is demonstrated by numerical experiments. \keywords{adiative heat transfer equations.}

\UDC{517.95}
\AMS{35K55, 35Q79}
\SupportedBy{}
\submitted{20 февраля 2018 г.}

\title{Граничная обратная задача для уравнений сложного теплообмена}

\author{П.\,Р.~Месенев}{Дальневосточный федеральный университет, 690950, г.~Владивосток, ул.~Суханова,~8;}{mesenev.pr@gmail.com}
\author{А.\,Ю.~Чеботарев}{Институт прикладной математики ДВО РАН, 690041, Владивосток, ул. Радио, 7}{cheb.iam@dvo.ru}

\makeface

\abstract Рассмотрена задача по восстановлению отражающих свойств участка границы для стационарной модели радиационно-кондуктивного теплообмена в трехмерной области. Приведен алгоритм решения задачи, эффективность которого проиллюстрирована численными примерами.

\keywords{уравнения радиационного теплообмена, квазирешение обратной задачи, метод градиентного спуска.}

\input{1_intro}
\input{2_state}
\input{3_formalization}
\input{4_analysis}
\input{5_experiments}

\bibliographystyle{unsrt}
\bibliography{bibliography}
\end{document}