### 2020-05-09 phd_3/article/article.tex
refactor + template for new experiment

\documentclass[12pt]{article}
\usepackage[russian]{babel}
\usepackage{amsmath}
\usepackage{amsfonts,amssymb}
\usepackage{amsbib}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{float}

\newtheorem{theorem}{Теорема}
\newtheorem{lemma}{Лемма}
\newtheorem{corollary}{Следствие}
\newtheorem{definition}{Определение}
\newtheorem{remark}{Замечание}
\newtheorem{problem}{Задача}

\renewcommand{\div}{\mathop{\mathrm{div}}\nolimits}

\begin{document}

    \section{Постановка задачи}\label{sec:постановка-задачи}
    \input{1_problem}

    \section{ЧИСЛЕННЫЕ ЭКСПЕРИМЕНТЫ}\label{sec:численные-эксперименты}
    \input{2_experiments}

\end{document}