### 2020-07-24 phd_3/article/article.tex

\documentclass[12pt,a4paper]{report}%fleqn,leqno
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{extsizes}
\usepackage[unicode, colorlinks=false, urlcolor=blue, pdfborder={0 0 0 [0 0]},bookmarks=false]{hyperref}
\usepackage{mathtext}%русские буквы в формулах
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[T2A]{fontenc}

\begin{document}
    \include{1_problem}
\end{document}