### 2022-01-15 phd_3/article/biblio.tex
pre-alpha ver. 1exp

%! Suppress = LineBreak
\begin{thebibliography}{999}

    \bibitem{Pinnau07}
    Pinnau R. Analysis of optimal boundary control for radiative heat transfer modeled by $SP_1$-system //
    Commun. Math. Sci. 2007. V. 5. \textnumero~4. P. 951--969.

    \bibitem{AMC-13}
    Kovtanyuk A.E., Chebotarev A.Yu. An iterative method for solving a complex heat transfer problem // Appl. Math. Comput. 2013. V. 219. \textnumero~17. P. 9356--9362.
    \bibitem{Kovt14-1}
    Kovtanyuk A.E., Chebotarev A.Yu., Botkin N.D., Hoffmann K.-H. The unique solvability of a complex 3D heat transfer problem // J. Math. Anal. Appl. 2014. V. 409. \textnumero~2. P. 808--815.
    \bibitem{JVM-14}
    Ковтанюк А.Е., Чеботарев А.Ю. Стационарная задача сложного теплообмена // Ж. вычисл. матем. и матем. физ. 2014. Т. 54. \textnumero~4. С. 711--719.
    \bibitem{Kovt14-2}
    Ковтанюк А.Е., Чеботарев А.Ю. Стационарная задача свободной конвекции с радиационным теплообменом // Дифференц. ур-ния. 2014. Т. 50. \textnumero~12. С. 1590--1597.
    \bibitem{Kovt14-3}
    Kovtanyuk A.E., Chebotarev A.Yu., Botkin N.D., Hoffmann K.-H. Theoretical analysis of an optimal control problem of conductive-convective-radiative heat transfer // J. Math. Anal.
    Appl. 2014. V. 412. \textnumero~1. P. 520--528.
    \bibitem{Grenkin1}
    Гренкин Г.В., Чеботарев А.Ю. Нестационарная задача сложного теплообмена // Ж. вычисл. матем. и матем. физ. 2014. Т. 54. \textnumero~11. С. 1806--1816.
    \bibitem{CNSNS-15}
    Kovtanyuk A.E., Chebotarev A.Yu., Botkin N.D., Hoffmann K.-H. Unique solvability of a steady-state complex heat transfer model // Commun. Nonlinear Sci. Numer. Simul. 2015. V. 20. \textnumero~3. P. 776--784.

    \bibitem{Grenkin3}
    Гренкин Г.В., Чеботарев А.Ю. Неоднородная нестационарная задача сложного теплообмена // Сиб. электрон. матем. изв. 2015. Т. 12. С. 562--576.
    \bibitem{Grenkin4}
    Гренкин Г.В., Чеботарев А.Ю. Нестационарная задача свободной конвекции с радиационным теплообменом // Ж. вычисл. матем. и матем. физ. 2016. Т. 56. \textnumero~2. С. 275--282.
    \bibitem{Grenkin5}
    Grenkin G.V., Chebotarev A.Yu., Kovtanyuk A.E., Botkin N.D., Hoffmann K.-H. Boundary optimal control problem of complex heat transfer model // J. Math. Anal. Appl. 2016. V. 433. \textnumero~2. P. 1243--1260.

    \bibitem{JMAA-16}
    Kovtanyuk A.E., Chebotarev A.Yu., Botkin N.D., Hoffmann K.-H. Optimal boundary control of a steady-state heat transfer model accounting for radiative effects // J. Math. Anal. Appl.
    2016. V. 439. \textnumero~2. P. 678--689.
    \bibitem{AMC-16}
    Chebotarev A.Yu., Kovtanyuk A.E., Grenkin G.V., Botkin N.D., Hoffmann K.-H\@.
    Nondegeneracy of optimality conditions in control problems for a radiative-conductive heat transfer model // Appl. Math. Comput. 2016. V. 289. P. 371--380.
    \bibitem{JVM-16}
    Ковтанюк А.Е., Чеботарев А.Ю. Нелокальная однозначная разрешимость стационарной задачи сложного теплообмена // Ж. вычисл. матем. и матем. физ. 2016. Т. 56. \textnumero~5. С. 816--823.
    \bibitem{ESAIM}
    Chebotarev A.Yu., Grenkin G.V., Kovtanyuk A.E. Inhomogeneous steady-state problem of complex heat transfer // ESAIM Math. Model. Numer. Anal. 2017. V. 51. \textnumero~6. P. 2511--2519.

    \bibitem{CNSNS18}
    Chebotarev, A.Y., Grenkin, G.V., Kovtanyuk, A.E., Botkin, N.D., Hoffmann, K.-H. Diffusion approximation of the radiative-conductive heat transfer model with Fresnel matching conditions// Communications in Nonlinear Science and Numerical Simulation. 57 (2018) 290-298.

    \bibitem{JMAA-18}
    Chebotarev A.Yu., Grenkin G.V., Kovtanyuk A.E., Botkin N.D., Hoffmann K.-H. Inverse problem with finite overdetermination for steady-state equations of radiative heat exchange // J. Math. Anal. Appl. 2018. V. 460. \textnumero~2. P. 737--744.
    \bibitem{JMAA-19}
    Chebotarev A.Yu., Pinnau R. An inverse problem for a quasi-static approximate model of radiative heat transfer // J. Math. Anal. Appl. 2019. V. 472. \textnumero~1. P. 314--327.
    \bibitem{JVM-19-INV} Гренкин Г.В., Чеботарев А.Ю. Обратная задача для уравнений сложного теплообмена // Журнал вычислительной математики и математической физики. 2019, том 59, N 8, с. 1420-1430.

    \bibitem{CNSNS19}
    Chebotarev A.Y., Kovtanyuk A.E., Botkin N.D. Problem of radiation heat exchange with boundary conditions of the Cauchy type // Communications in Nonlinear Science and Numerical Simulation. 75 (2019) 262-269.

    \bibitem{CMMP20}
    Колобов А.Г., Пак Т.В., Чеботарев А.Ю.
    Стационарная задача радиационного теплообмена с граничными условиями типа Коши
    // Журнал вычислительной математики и математической физики.
    2019, том 59, N 7, с. 1258--1263.

    \bibitem{Amosov05}
    Амосов~А.А. Глобальная разрешимость одной нелинейной
    нестационарной задачи с нелокальным краевым условием типа
    теплообмена излучением~// Дифференц. ур-ния. Т.~41. N~1.
    2005. С.~93--104.

    \bibitem{Amosov10}
    Amosov~A.A. Stationary nonlinear nonlocal problem of
    radiative-conductive heat transfer in a system of opaque bodies
    with properties depending on the radiation frequency~// J\@.
    Math. Sc. 2010. V.~164. N~3. P.~309--344.

    \bibitem{Amosov16}
    Amosov A. Unique Solvability of a Nonstationary Problem of Radiative - Conductive
    Heat Exchange in a System of Semitransparent Bodies // Russian J. of Math.
    Phys. 2016. V. 23, N~3. P.~309-334.

    \bibitem{Amosov17}
    Amosov A.A. Unique Solvability of Stationary Radiative - Conductive Heat Transfer
    Problem in a System of Semitransparent Bodies // J. of Math. Sc.(United
    States). 2017. V. 224. N~5. P.~618-646.

    \bibitem{Amosov20-1}
    Amosov, A.A. Asymptotic Behavior of a Solution to the Radiative Transfer Equation in a Multilayered Medium with Diffuse Reflection and Refraction Conditions // J Math Sci 244, 541-575 (2020).

    \bibitem{Amosov20}
    Amosov A.A., Krymov N.E. On a Nonstandard Boundary Value Problem Arising in Homogenization of Complex Heat Transfer Problems // J. of Math. Sc.(United
    States). 2020. V. 244. P.~357 - 377.

    \bibitem{Kufner} S.~Fu\v{c}ik, A.~Kufner, Nonlinear differential equations,
    Elsevier, Amsterdam--Oxford--New York, 1980.

    \bibitem{10} A.V.~Fursikov, Optimal Control of Distributed
    Systems. Theory and Applications, American Math. Soc., 2000.

    \bibitem{11} A.D.~Ioffe, V.M.~Tikhomirov, Theory of Extremal
    Problems, North-Holland, Amsterdam, 1979.

    \bibitem{fenics} M. S. Alnaes, J. Blechta, J. Hake, A. Johansson,
    B. Kehlet, A. Logg, C. Richardson, J. Ring, M. E. Rognes, G. N. Wells
    The FEniCS Project Version 1.5
    Archive of Numerical Software, vol. 3, 2015,

    \bibitem{dolfin} A. Logg and G. N. Wells, DOLFIN: Automated Finite Element Computing
    ACM Transactions on Mathematical Software, vol. 37, 2010

    \bibitem{mesenev-github} \url{https://github.com/mesenev/articles_src}

\end{thebibliography}