### 2022-02-05 phd_3/article/article.tex

\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\usepackage{extsizes}
\usepackage{mathtext}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{float}
\usepackage{subfig}
\usepackage[T2A]{fontenc}

\begin{document}
    \include{1_problem}
    \include{2_formalization}
    \include{4_optimality}
    \include{5_experiments}
    \include{biblio}
\end{document}